<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An Introduction to the soundecology package}
-->

# An Introduction to soundecology package #

## Introduction ##

The  **soundecology** package has functions to calculate indices for soundscape ecology and other ecology research that uses audio recordings.


## Indices ##

The package **soundecology** can calculate several indices:

 - Acoustic Complexity Index (ACI) from Pieretti *et al.* 2011
 - Normalized Difference Soundscape Index (NDSI) from REAL (http://www.real.msu.edu) and Kasten *et al.* 2012
 - Boiacoustic Index from Boelman *et al.* 2007.
 - Acoustic Diversity Index (ADI) from Villanueva-Rivera *et al.* 2011

The examples below use two sound recordings to illustrate the use of the functions:
 - A Wave object included in the package (tropicalsound)
 - A wav file available on the web

### Acoustic Complexity Index (ACI) ###

The ACI is based on the "observation that many biotic sounds, such as bird songs, are 
characterized by an intrinsic variability of intensities, while some types
of human generated noise (such as car passing or airplane transit)
present very constant intensity values" (Pieretti, et al. 2011).

To calculate the ACI of the example Wave object **tropicalsound**, use these commands:
```{r results='show', message=FALSE, warning=FALSE}
#Load the package
require(soundecology)

#Call the Wave object into memory
data(tropicalsound)

#Run the function
acoustic_complexity(tropicalsound)
```

Example with wav file:
```{r results='show', message=FALSE, warning=FALSE}
#Download a wave file from the web
download.file("http://1159sequoia05.fnr.purdue.edu/primer/sounds/SM87_20080420_000000.wav", destfile="SM87_20080420_000000.wav")

#Load file as an object called soundfile
soundfile <- readWave("SM87_20080420_000000.wav")

#Delete the downloaded wave file
unlink("SM87_20080420_000000.wav")

#Run the function on this object and save the results in a new variable called soundfile.aci
soundfile.aci <- acoustic_complexity(soundfile)

#Print the ACI value for that file, stored in the soundfile.aci variable
print(soundfile.aci$AciTotAll_left)
```




### Normalized Difference Soundscape Index (NDSI) ###

The Normalized Difference Soundscape Index (NDSI), from REAL (http://www.real.msu.edu) and Kasten, *et al.* 2012, 
seeks to "estimate the level of anthropogenic disturbance on the soundscape by
computing the ratio of human-generated (anthrophony) to biological
(biophony) acoustic components found in field collected sound samples".

```{r results='show', message=FALSE, warning=FALSE}
library(soundecology)
data(tropicalsound)
NDSI <- ndsi(tropicalsound)
print(NDSI$ndsi_left)

summary(NDSI)
```

### Bioacoustic Index ###

The Bioacoustic Index, from Boelman *et al.* 2007, is calculated as the "area under each curve included all frequency bands associated
with the dB value that was greater than the minimum dB value for each curve. The area values are thus a
function of both the sound level and the number of frequency bands used by the avifauna".

```{r results='show', message=FALSE, warning=FALSE}
library(soundecology)
data(tropicalsound)
bioindex <- bioacoustic_index(tropicalsound)
print(bioindex$left_area)

summary(bioindex)
```
